---
title: "mustlinkMix"
author: "Ultán P. Doherty"
date: "2024-05-08"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install `mustlinkMix`.
```{r, eval=FALSE}
remotes::install_github("UltanPDoherty/mustlinkMix")
```

## Load and plot data from the `healthyFlowData` package.
```{r hfd1_setup, message=FALSE, warning=FALSE}
library(healthyFlowData)
data(hd)
hfd1 <- hd.flowSet[[1]]@exprs

GGally::ggpairs(hfd1, upper = list(continuous = "density"), progress = FALSE)
```

## Prepare a plusminus table which describes three populations.

* CD4+ T Cells (CD4+CD8-CD3+CD19-)
* CD8+ T Cells (CD4-CD8+CD3+CD19-)
* B Cells      (CD4-CD8-CD3-CD19+)

```{r plusminus, message=FALSE, warning=FALSE}
plusminus1 <- as.data.frame(rbind(
  "CD4+_T" = c(+1, -1, +1, -1),
  "CD8+_T" = c(-1, +1, +1, -1),
  "B"      = c(-1, -1, -1, +1)
))
colnames(plusminus1) <- colnames(hfd1)
plusminus1
```

## Excel can be used to save or create tables (`openxlsx` package).
```{r openxlsx}
openxlsx::write.xlsx(
  plusminus1,
  "~/plusminus.xlsx",
  rowNames = TRUE,
  colNames = TRUE
)

plusminus2 <- openxlsx::read.xlsx(
  "~/plusminus.xlsx",
  rowNames = TRUE,
  colNames = TRUE
)
```

## Use the `gatetree` function from the `gateTree` package.
```{r gatetree, message=FALSE, warning=FALSE}
hfd1_gatetree <- gateTree::gatetree(
  hfd1,
  plusminus2,
  min_scaled_bic_diff = 50,
  min_depth = 10,
  show_plot = c(FALSE, FALSE)
)
```

## Plot the data, coloured according to the `gateTree` labels.
```{r ggpairs_gatetree, message=FALSE, warning=FALSE}
GGally::ggpairs(hfd1,
  progress = FALSE,
  upper = list(continuous = "density"),
  ggplot2::aes(colour = as.factor(1 + hfd1_gatetree$labels))
) +
  ggokabeito::scale_colour_okabe_ito(order = c(9, 1, 2, 3)) +
  ggokabeito::scale_fill_okabe_ito(order = c(9, 1, 2, 3))
```

## Apply mustlink.
```{r tree_plot, message=FALSE, warning=FALSE}
hfd1_mustlink <- mustlinkMix::mustlink(
  hfd1,
  clust_num = 5,
  zone_matrix = hfd1_gatetree$subsetter,
  zone_percent = 100,
  init_seed = 123,
  init_method = "mlkmpp"
)
```

## Plot the data, coloured according to the mustlink labels.
```{r ggpairs_mustlink, message=FALSE, warning=FALSE}
GGally::ggpairs(hfd1,
  progress = FALSE,
  upper = list(continuous = "density"),
  ggplot2::aes(colour = as.factor(hfd1_mustlink$clust_labels))
) +
  ggokabeito::scale_colour_okabe_ito(order = c(1, 2, 3, 5, 6)) +
  ggokabeito::scale_fill_okabe_ito(order = c(1, 2, 3, 5, 6))
```
